---
title: "Aides Drug"
author: "Wilhelmus Elersich"
date: "9/22/22"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
H_0: There is no statistical relationship between treatment and contracting AIDS.

H_1: Receiving the drug triuMpsal reduces the likelyhood of contracting AIDS.

First, looking at the data:

```{R}
dim(triUmp)
names(triUmp)
```
Here we are explaining a categorical variable by a another, distinct, categorical variable. We should thus use methods listed in the cat_cat file to determine the relationship between these variables. We should remove all cases who are not drug users, and then only select the variables needed to answer our question. We should also convert all character variables to factor variables, to make working with these variables easier. The helper file MyTricks contains a code snippet that can be modified to do exactly what we want.

```{r}
ABadTime <- filter(triUmp, DrugUser == "yes") %>%
  select(X, treatment, infected) %>%
  mutate(across(where(is.character), as.factor)) %>%
  arrange(desc(infected)) 
head(ABadTime, 10)
```  
To analyze the results, we should use a graphical method, and a numerical method, both from Cat_Cat. The graphical method we will use is the boxplot, and the numerical method is the "Numeric Summaries" snippet from Cat_Cat, which will give two tables which show the percentages of our target population who fall into the categories we are trying to find a relationship between. Finally, we should conduct the Chi-squared test and the fisher test, both of which can be found in Cat_Cat, to determine the likelyhood of the null hypothesis being true.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected + treatment,data=ABadTime)
barchartGC(~infected + treatment,data=ABadTime, type="percent")

```

According to these results, nearly 80% of those who had contracted AIDS had only received placebo, while those who took the drug only make up little more than 20% of those who contracted AIDS. This would suggest that the drug is effective for drug users.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment,data=ABadTime)
rowPerc(table1)
colPerc(table1)
```
Just like in our graphical results, we can see that 24.29% of those who contracted AIDS had taken the drug, while 75.51% of those who had contracted AIDS were on placebo. Furthermore, 2.89% of those who had taken the drug contracted AIDS, which is less than half of the percentage of those who took placebo who contracted AIDS, at 7.77%. So long as the null hypothesis is false, we can confidently say that the drug reduces the probability of contracting AIDS for drug users.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-values from these tests are extremely low, so we can confidently reject the null hypothesis. Furthermore, the odds ratio, 0.3534128, quantifies what has been alluded to in the graphical and numerical results, that a drug user who took the drug is only 35.4128% as likely to contract AIDS as a drug user who had taken placebo.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

From the above results we can conclude that the drug triUmp significantly lowers the chances of contracting AIDS for drug users; those drug users who take the drug are only 35.4128% as likely to contract AIDS as those who had taken placebo. Therefore, if a drug user wishes to avoid contracting AIDS, that drug user should take the drug. However, the drug might not work for everyone, since these results only apply to a generalized group of drug users; other factors might affect the drug's efficacy.
